---
title: "Case Study 1: Lifespans of World Leaders"
author: "Ethan Shen, Malavi Ravindran, Steven Herrera Tenorio, Anna Darwish"
fontsize: 12pt
geometry: "left=1.3cm,right=1.3cm,top=1.4cm,bottom=1.4cm"
output: 
  pdf_document:
     latex_engine: xelatex
     number_sections: true
---
```{r, function checking for installed packages, include=FALSE}
# Validate that all necessary packaged have been downloaded, install otherwise or throw err package DNE
pkgTest <- function(x)
{
  if (!require(x,character.only = TRUE))
  {
    install.packages(x,repos = "http://cran.r-project.org", dep=TRUE)
    if(!require(x,character.only = TRUE)) stop("Package not found")
  }
}
```

```{r include=FALSE}
# Installing packages 
# Additionally, if you are struggling to download RJags, please visit the following link for help. It can also help to try to download the package locally or in RStudio Cloud: https://sites.google.com/a/utexas.edu/edm-principalstratification/downloading-installing-r-jags-rstudio
pkgTest("tidyverse")
pkgTest("knitr")
pkgTest("kableExtra")
pkgTest("stringr")
pkgTest("coda")
pkgTest("tidybayes")
pkgTest("ggdist")
pkgTest("gridExtra")
pkgTest("grid")
pkgTest("R2jags") 
pkgTest("readr")
pkgTest("dplyr")
pkgTest("ggplot2")
pkgTest("lubridate")
pkgTest("tidyr")
pkgTest("tibble")
```

```{r include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(stringr)
library(coda)
library(tidybayes)
library(ggdist)
library(gridExtra)
library(grid)
library(R2jags)
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(tibble)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
ggplot2::theme_set(new = theme_bw())
```

```{r Load & Clean Data}
## The following code updates the dataset to ensure timestamps reflect the date of this study, 8/31/2020
load("Leaders.Rdata")
leaders <- as.data.frame(leaders)
#head(leaders)
prev_date <- as.Date("31/07/2020",format="%d/%m/%Y")
current_date <- as.Date("31/08/2020",format="%d/%m/%Y")

## remove days string and convert to numeric
leaders <- leaders %>% 
  mutate(Age.Event = sub(".*?(\\d+\\.*\\d*).*", "\\1", Age.Event)) %>% data.frame
leaders$Age.Event <- as.numeric(leaders$Age.Event)

## update age event by adding 1/12 to leaders still alive (none died in past month so easy to do case when)
leaders <- leaders %>% 
  mutate(Age.Event = case_when(Event.Date == prev_date ~ Age.Event + 1/12,
                               TRUE ~ Age.Event)) %>% data.frame

## update event dates to be current day since none of the leaders in this set died in the past month
leaders <- leaders %>% 
  mutate(Event.Date = case_when(Event.Date == prev_date ~ as.Date(current_date,format="%d/%m/%Y"),
                                TRUE ~ Event.Date)) %>% data.frame

## update Emperor Jianwen
# leaders <- leaders %>% 
#   mutate(Fail = case_when(Name == "Jianwen Emperor" ~ 1,
#                           TRUE ~ Fail),
#          Censored = case_when(Name == "Jianwen Emperor" ~ 0,
#                               TRUE ~ Censored)) %>% data.frame
# 
# ## multiplying time by 10 
# leaders <- leaders %>% 
#   mutate(decade = time * 10)
alive = c("Barack Obama","Francis","14th Dalai Lama (Tenzin Gyatso)", "Naruhito (Emperor Reiwa)")
filtered_leaders = leaders %>% 
  filter(!Name %in% alive)

#load("leaders_model_r_25_2e-1_beta_1.Rdata")
#Bayesian_Leaders_output <- rjags_model
```

```{r Covariates from Filtered Dataset}
## The following code sets up the covariates and censoring information for the soon to-be-created models, after removing the ones we
## intend to make cast predictions for
alive = c("Barack Obama","Francis","14th Dalai Lama (Tenzin Gyatso)", "Naruhito (Emperor Reiwa)")
filtered_leaders = leaders %>% 
  filter(!Name %in% alive)

n <- nrow(filtered_leaders)
Age.Event <- filtered_leaders$Age.Event
censored <- filtered_leaders$Censored
survival <- ifelse(censored == 0, Age.Event, NA)
censoring_limits <- ifelse(censored == 0, 100, Age.Event)

# Co-variates
x_1 <- filtered_leaders$time - mean(filtered_leaders$time)
x_2 <- filtered_leaders$Type %>% as.factor() %>% relevel("Pope")
```

```{r Predictive Information}
# predictive info for Francis, Obama, and Naruhito
# 1 = Pope, 2 = ChineseEmp, 3 = DalaiLama, 4 = JapanEmp, 5 = UsPres
#Dalai Lama
time_DL <- leaders[leaders$Name == "14th Dalai Lama (Tenzin Gyatso)",]$time - mean(filtered_leaders$time)
type_DL <- 3
present_age_DL <- leaders[leaders$Name == "14th Dalai Lama (Tenzin Gyatso)",]$Age.Event


#Francis
time_Francis <- leaders[leaders$Name == "Francis",]$time - mean(filtered_leaders$time)
type_Francis <- 0
present_age_Francis <- leaders[leaders$Name == "Francis",]$Age.Event

#Obama
time_Obama <- leaders[leaders$Name == "Barack Obama",]$time - mean(filtered_leaders$time)
type_Obama <- 5
present_age_Obama <- leaders[leaders$Name == "Barack Obama",]$Age.Event

#Naruhito
time_Naruhito <- leaders[leaders$Name == "Naruhito (Emperor Reiwa)",]$time - mean(filtered_leaders$time)
type_Naruhito <- 4
present_age_Naruhito <- leaders[leaders$Name == "Naruhito (Emperor Reiwa)",]$Age.Event
```

```{r}
## external validation - Churchill + Sun Yat-sen
## assuming Churchill is similar to US President and Sun is similar to Chinese emperor, 
## so Churchill's type is 5 and Sun's type is 2
cent_date <- "1300-01-01" %>% as.Date()

birthdate_WC <- "1874-11-30" %>% as.Date()
time_WC <- (decimal_date(birthdate_WC) - decimal_date(cent_date)) / 100
type_WC <- 5
time_WC_cent <- time_WC - mean(filtered_leaders$time)

birthdate_sun <- "1866-11-12" %>% as.Date()
time_sun <- (decimal_date(birthdate_sun) - decimal_date(cent_date)) / 100
type_sun <- 2
time_sun_cent <- time_sun - mean(filtered_leaders$time)
```

```{r}
## censoring limits
present_age <- ifelse(censored == 0, 0, filtered_leaders$Age.Event)
upper_age <- 110
```

```{r Stored Variables in Models}
## The following code stores the list of variables we wish to store in the model Rdata files from the RJAGS code
data_leaders <- list("n",
                     "censored",
                     "survival",
                     "censoring_limits",
                     "x_1",
                     "x_2",
                     "present_age",
                     "upper_age",
                     "time_Francis",
                     "present_age_Francis",
                     "time_Obama",
                     "type_Obama",
                     "present_age_Obama",
                     "time_DL",
                     "type_DL",
                     "present_age_DL",
                     "time_Naruhito",
                     "type_Naruhito",
                     "present_age_Naruhito",
                     "time_WC_cent", 
                     "type_WC",
                     "time_sun_cent",
                     "type_sun"
)
```

```{r Generate Model Text}
#RJAGs allows users to input either the text of the model they hope to create (where information such as likelihood, priors, and predictive information instructions are necessary) or input an R function. To minimize code duplication, the following method concatenates the likelihood information, prior specification, and predictive information. Anyone who wishes to build off of these results can input the priors (see sections Model 1a - Model 6b for various examples).

generate_model_text = function(priors=""){
  model_file <- tempfile()
  
  likelihood = "#Likelihood
  for (i in 1:n){ 
    censored[i] ~ dinterval(survival[i], censoring_limits[i]) 
    survival[i] ~ dweib(r, mu[i]) # Basic Weibull assumption
    mu[i] <- exp(beta[i])  # Defining beta as log(mu)
    beta[i] <- beta_0 + beta_1*x_1[i] + beta_type[x_2[i]] + beta_int[x_2[i]] * x_1[i] 
    y_pred_censored[i] ~ dweib(r, mu[i])T(present_age[i], upper_age)
  }
    
  #Priors
  beta_type[1] <- 0
  beta_int[1] <- 0
  "
  
  predicted_survivals = "
  #Define the alphas
  alpha_0 <- - beta_0 / r
  alpha_1 <- - beta_1 / r
  percentage_increase_birthyr <- 100*(exp(alpha_1) - 1)
  for (i in 2:5) {
    alpha_type[i] <- - beta_type[i] / r
    alpha_int[i] <- -beta_int[i] / r
    percentage_increase_type[i] <- 100*(exp(alpha_type[i]) - 1)
    percentage_increase_int[i] <- 100*(exp(alpha_int[i]) - 1)
  }
  
  beta_Francis <- beta_0 + beta_1*time_Francis 
  mu_Francis <- exp(beta_Francis)
  survival_Francis ~ dweib(r, mu_Francis)T(present_age_Francis, upper_age)
  
  beta_Obama <- beta_0 + beta_1*time_Obama + 
    beta_type[type_Obama] + beta_int[type_Obama]*time_Obama
  mu_Obama <- exp(beta_Obama)
  survival_Obama ~ dweib(r, mu_Obama)T(present_age_Obama, upper_age)
  
  beta_Naruhito <- beta_0 + beta_1*time_Naruhito + beta_type[type_Naruhito] +
    beta_int[type_Naruhito]*time_Naruhito
  mu_Naruhito <- exp(beta_Naruhito)
  survival_Naruhito ~ dweib(r, mu_Naruhito)T(present_age_Naruhito, upper_age)
  
  beta_DL <- beta_0 + beta_1*time_DL + beta_type[type_DL] + beta_int[type_DL]*time_DL
  mu_DL <- exp(beta_DL)
  survival_DL ~ dweib(r, mu_DL)T(present_age_DL, upper_age)
  
  beta_WC <- beta_0 + beta_1*time_WC_cent + beta_type[type_WC] + 
    beta_int[type_WC]*time_WC_cent
  mu_WC <- exp(beta_WC)
  survival_WC ~ dweib(r, mu_WC)T(0, upper_age)
  
  beta_sun <- beta_0 + beta_1*time_sun_cent + beta_type[type_sun] + 
    beta_int[type_sun]*time_sun_cent
  mu_sun <- exp(beta_sun)
  survival_sun ~ dweib(r, mu_sun)T(0, upper_age)
  "
  
  model_text <- paste("model{ ",likelihood, priors, predicted_survivals, "}")
  writeLines(model_text,con=model_file)
  return(model_file)
}
```

```{r Generate or Load Model}
#The following utility function takes in the priors and name of the model file the user intends to create before running RJAGS. In the case the model_name exists in the current working directory, this function will instead load in that model and return it for further exploration.
generate_model = function(priors,model_name,num_iter = 250000){
  model_file <- generate_model_text(priors)
  if (file.exists(model_name)) {
    load(file = model_name)
  } 
  else {
    rjags_model <-  jags.parallel(data = data_leaders,  
                                  parameters.to.save = c("beta_0",
                                                         "beta_1",
                                                         "beta_type",
                                                         "beta_int",
                                                         "r",
                                                         "y_pred_censored",
                                                         "survival_Francis",
                                                         "survival_Obama",
                                                         "survival_Naruhito",
                                                         "survival_DL",
                                                         "survival_WC",
                                                         "survival_sun",
                                                         "alpha_0",
                                                         "alpha_1",
                                                         "percentage_increase_birthyr",
                                                         "alpha_type",
                                                         "alpha_int",
                                                         "percentage_increase_type",
                                                         "percentage_increase_int",
                                                         "survival"
                                  ), 
                                  n.iter = num_iter, #increase for better model output - 250K
                                  n.chains = 3,
                                  model.file = model_file)
    save(rjags_model, file = model_name)
  }
  return(rjags_model)
}
```

```{r}
## Model From Paper, Original Id: 5a
priors <- "
  # prior on leadership type coefficients (notice there is no need to specify a prior for the baseline)
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 1)
    beta_int[i] ~ dnorm(0.0, 1)
  }
  
  # prior on intercept
  beta_0 ~ dnorm(0.0, 0.01)
  
  # prior on time variable
  beta_1 ~ dnorm(0.0, 1)
  
   # prior on r 
  r ~ dgamma(10,0.5)
  "
Bayesian_Leaders_output <- generate_model(priors,"leaders_model_r_10_5e-1_beta_1.Rdata",500000)
```


```{r}
## Model 1a: r ~ exp(0.001), Betas ~ N(0,1)
priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 1)
    beta_int[i] ~ dnorm(0.0, 1)
  }
  
  beta_0 ~ dnorm(0.0, 0.01) # Prior on beta_0 is normal with low precision
  beta_1 ~ dnorm(0.0, 1) # Prior on beta_1 is normal with low precision
  
  r ~ dexp(0.001) # Prior on r
  "
Bayesian_Leaders_V1a <- generate_model(priors,"leaders_model_r_1e-3_beta_1.Rdata")
```

```{r}
## Model 1b: r ~ exp(0.001), Betas ~ N(0,100)
priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, .01)
    beta_int[i] ~ dnorm(0.0, .01)
  }
  
  beta_0 ~ dnorm(0.0, 0.01) # Prior on beta_0 is normal with low precision
  beta_1 ~ dnorm(0.0, .01) # Prior on beta_1 is normal with low precision
  
  r ~ dexp(0.001) # Prior on r
  "
Bayesian_Leaders_V1b <- generate_model(priors,"leaders_model_r_1e-3_beta_100.Rdata")
```

```{r}
## Model 2a - r ~ exp(0.1), Betas ~ N(0,1)
priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 1)
    beta_int[i] ~ dnorm(0.0, 1)
  }
  
  beta_0 ~ dnorm(0.0, 0.01)
  beta_1 ~ dnorm(0.0, 1)
  
  r ~ dexp(0.1) # Prior on r
  "
Bayesian_Leaders_V2a <- generate_model(priors,"leaders_model_r_1e-1_beta_1.Rdata")
```

```{r}
## Model 2b - r ~ exp(0.1), Betas ~ N(0,100)

priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, .01)
    beta_int[i] ~ dnorm(0.0, .01)
  }
  
  beta_0 ~ dnorm(0.0, 0.01)
  beta_1 ~ dnorm(0.0, .01)
  
  r ~ dexp(0.1) # Prior on r
  "
Bayesian_Leaders_V2b <- generate_model(priors,"leaders_model_r_1e-1_beta_100.Rdata")
```

```{r}
## Model 3a - r ~ exp(1), Betas ~ N(0,1)

priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 1)
    beta_int[i] ~ dnorm(0.0, 1)
  }
  
  beta_0 ~ dnorm(0.0, 0.01)
  beta_1 ~ dnorm(0.0, 1)
  
  r ~ dexp(1) # Prior on r
  "
Bayesian_Leaders_V3a <- generate_model(priors,"leaders_model_r_1_beta_1.Rdata")
```

```{r}
## Model 3b - r ~ exp(1), Betas ~ N(0,100)
priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 100)
    beta_int[i] ~ dnorm(0.0, 100)
  }
  
  beta_0 ~ dnorm(0.0, 0.01)
  beta_1 ~ dnorm(0.0, .01)
  
  r ~ dexp(1) # Prior on r
  "
Bayesian_Leaders_V3b <- generate_model(priors,"leaders_model_r_1_beta_100.Rdata")
```

```{r}
## Model 4a - r ~ exp(2), Betas ~ N(0,1)
priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 1)
    beta_int[i] ~ dnorm(0.0, 1)
  }
  
  beta_0 ~ dnorm(0.0, 0.01)
  beta_1 ~ dnorm(0.0, 1)
  
  r ~ dexp(2) # Prior on r
  "
Bayesian_Leaders_V4a <- generate_model(priors,"leaders_model_r_2_beta_1.Rdata")
```



```{r}
## Model 4b - r ~ exp(2), Betas ~ N(0,100)
priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 0.01)
    beta_int[i] ~ dnorm(0.0, 0.01)
  }
  
  beta_0 ~ dnorm(0.0, 0.01)
  beta_1 ~ dnorm(0.0, 0.01)
  
  r ~ dexp(2) # Prior on r
  "
Bayesian_Leaders_V4b <- generate_model(priors,"leaders_model_r_2_beta_100.Rdata")
```


```{r}
## Model 5b - r ~ gamma(10,0.5), Betas ~ N(0,100)
priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 0.01)
    beta_int[i] ~ dnorm(0.0, 0.01)
  }
  
  beta_0 ~ dnorm(0.0, 0.01)
  beta_1 ~ dnorm(0.0, 0.01)
  
  r ~ dgamma(10,0.5) # Prior on r
  "
Bayesian_Leaders_V5b <- generate_model(priors,"leaders_model_r_10_5e-1_beta_100.Rdata")
```

```{r}
## Model 6a - r ~ gamma(25,0.2), Betas ~ N(0,1)
priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 1)
    beta_int[i] ~ dnorm(0.0, 1)
  }
  
  beta_0 ~ dnorm(0.0, 0.01)
  beta_1 ~ dnorm(0.0, 1)
  
  r ~ dgamma(25,0.2) # Prior on r
  "
Bayesian_Leaders_V6a <- generate_model(priors,"leaders_model_r_25_2e-1_beta_1.Rdata")
```

```{r}
## Model 6b - r=gamma(25,0.2), Betas: N(0,100)
priors <- "
  for(i in 2:5) {
    beta_type[i] ~ dnorm(0.0, 0.01)
    beta_int[i] ~ dnorm(0.0, 0.01)
  }

  beta_0 ~ dnorm(0.0, 0.01)
  beta_1 ~ dnorm(0.0, 0.01)

  r ~ dgamma(25,0.2) # Prior on r
  "
Bayesian_Leaders_V6b <- generate_model(priors,"leaders_model_r_25_2e-1_beta_100.Rdata")
```

# Introduction 

World leaders have measurable influence in politics, even beyond their election terms. For this reason, it is critical to explore their lifespans. Extending the investigation from Stander et al.[[1]][Bibliography], we analyze the survival times of 177 world leaders, including Roman Catholic Popes, U.S. Presidents, Chinese and Japanese Emperors, and Dalai Lamas. We propose a Weibull distribution for modeling the survival time of these world leaders and provide a mechanism for predicting their survival times within a Bayesian framework. 

Our goal is to ultimately determine how year of birth and type of leadership are related to a leader's survival time. Furthermore, we seek to understand if the relationship between year of birth and lifespan is dependent on type of leadership. Finally, we provide point and interval predictions of the lifespans of living leaders, along with posterior predictive distributions to make comparative statements about future survival times. 

# Data 

Leaders in this dataset date from the 14th century through present day. Survival time, in years, is defined to be either the leader’s current age, if alive, or age of death, if deceased. The 11 leaders that are currently alive or do not have a recorded death date are considered right censored observations - that is, we do not know their age of death, but know their current age. 10 leaders are currently alive, and Chinese Emperor Jiawen, who was born in the 14th century, disappeared and is presumably long dead by now. Additionally, we removed Pope Francis, the 14th Dalai Lama, Japanese Emperor Naruhito, and President Obama from our dataset, since we will be performing posterior predictions on their lifespans. 

Covariates of interest in this study are $type$ and $time$. $Type$ is a categorical variable corresponding to 5 different groups of world leaders. Of the $type$ levels, popes are the most represented with 63 observations while Dalai Lamas comprise the smallest group, with only 14 instances. The $time$ covariate represents a leader's year of birth, measured as the number of centuries since 1300. For example, as Pope Francis was born roughly 637 years after January 1, 1300, his $time$ value is represented as 6.37. We opted for this shifted and scaled version of birth year in order to prevent larger year values (1961, for example) from causing model instability. 

Below we will present some preliminary visualizations regarding our covariates of interest and survival outcome. 

## Exploratory Data Analysis 

```{r fig.align='center', fig.width=12, fig.height=3.75}
# The following code generates boxplots of lifespan across leadership types along with a scatter plot showing the trends in lifespan over time
gridExtra::grid.arrange(ggplot(leaders, aes(x = Type, y = Age.Event)) + 
                          geom_boxplot() + 
                          labs(title = "Lifespans vs Type of Leader", 
                               y="Lifespan (yrs)", x="Leader Type",
                               caption="Figure 1") + 
                          theme(
                            plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 18))
                        ,
                        ggplot(leaders, aes(x=Birth.Date, y=Age.Event, shape=Type, color=Type)) +
                          geom_point() + 
                          labs(title="Lifespans vs Birth Date",
                               caption="Figure 2") + 
                          xlab("Transformed Birth Date") +
                          geom_smooth(method = "lm", se=FALSE) + 
                          theme(axis.title.y = element_blank(),
                                plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 18)),
                        ncol=2)

```

Figure 1 displays lifespan across different leadership types. Among the Chinese Emperors, Japanese Emperors, and Dalai Lamas, the interquartile ranges of lifespan largely overlap. Popes and U.S. Presidents appear to have a slightly higher median lifespan. This is consistent with the fact that Popes and Presidents are elected much later in their lives, while Dalai Lamas and Chinese and Japanese Emperors may come to power when they are younger (either due to heredity or identification at a young age). An American citizen who dies in her youth, for example, can never become President due to the age threshold, so her short survival time can never be reflected in this dataset.

Figure 2 displays the change in lifespan across time for different types of leadership. It appears that lifespan has increased at similar rates across Chinese and Japanese Emperors, U.S. Presidents and Popes over time. However, for Dalai Lamas, the plot indicates that lifespan has decreased over time. This unexpected trend can be attributed to the untimely demise of several Dalai Lamas in the 19th century, with causes of death ranging from common illnesses to ceiling collapses [[5]][Bibliography]. 

# Methodology

## Model Selection 

We propose a Weibull distribution to model the survival time $T_i$ for the $i$th leader. This distribution is a generalization of the exponential distribution and can be used to model survival outcomes. The Weibull probability density function for $T_i$ takes the following form:

$$
f_{T_i}(t_i|r, \mu_i) = r\ \mu_i\ t_i^{r-1}\ exp(-\mu_i\ {t_i}^r),\ \ \ \ \ r,\  \mu_i, \ t_i >0
$$

The first parameter $r$ of the Weibull distribution is a positive scale parameter, and is constant across all leaders. The log of the second parameter, $\mu_i$, is a linear function of $time$, $type$, and the interaction effect between the previous two covariates. For ease of interpretability, $time$ was mean-centered.

Our proposed model is as follows: 

$$
\begin{aligned}
T_i \sim & \,Weibull(r, \mu_i)  \\
log(\mu_i) = \beta_0 + \beta_1\ time_i + \sum_{j=2}^{5} \beta_{2j} & \ I(type_i = j) + \sum_{j=2}^{5}  \beta_{3j}\ time_i * I(type_i = j)\ , &(1)
\end{aligned}
$$

In Model 1, $i$ corresponds to each world leader and $T_i$ refers to their survival time. The $time_i$ covariate represents birth year, measured in the number of centuries after 1300 that each leader was born, and $type_i$ refers to the individual's type of leadership. The reference level for the $type$ covariate is Pope, as it forms the largest category ($n=63$). The remaining levels $j = 2, 3, 4, 5$ correspond to Chinese Emperor, Dalai Lama, Japanese Emperor, and U.S. President respectively. The third term in the linear combination of covariates represents an interaction effect between $type$ and $time$. This interaction is used to account for the varying relationships between survival time and birth year that may exist across leadership types. 

To account for censored observations, we sample $T_i$ from a truncated distribution

$$
T_i \sim  \,Weibull(r,\, \mu_i)\,T(present\_age_i, \, upper\_age)
$$

Here, $present\_age_i$ is 0 for a leader who is deceased, and their current age if they are still alive. For deceased leaders, the lower bound of 0 prevents the model from making assumptions about their lifespan. For leaders who are currently alive, it ensures the model does not sample a survival time that is less than their current age. We set the $upper\_age$ to be 110. We selected this upper bound due to the fact that the United Nations predicts that only 0.0074% of the world's population will celebrate their 100th birthday and become a centenarian [[2]][Bibliography]. Of this small percentage of people, the Gerontology Research Group estimates only 0.15% to 0.25% will reach the age of 110 [[3]][Bibliography]. 

## Choice of Prior Distributions

We adopted the following prior distributions for $r,\ \beta_0, \  \beta_1,\  \beta_{2j}$ and $\beta_{3j}$:

$$
\begin{aligned}
&\,\,\,\,\,\,\,\,\,\,\,\,r \sim  \Gamma(10, 0.5)\\ 
&\,\,\,\,\,\,\,\,\,\beta_0\sim N(0, 100)\\ 
\beta_1,\ \beta_{2j}&,\  \beta_{3j} \sim N(0, 1), \  j = 2, 3, 4, 5
\end{aligned}
$$


Though there does not exist a strong body of literature to inform these priors, each of our choices resulted from research and consideration of context. For our prior on $\beta_0$, the intercept term, we chose a noninformative $N(0,100)$. As we do not believe the intercept is centered at 0, we imposed a high variance on the prior to allow it to explore the space. For the prior on $\beta_1$, the coefficient corresponding to the $time$ covariate, we decided to use a $N(0, 1)$ distribution. We were motivated to explore this zero-centered, low variance prior due to a study [[4]][Bibliography] citing that lifespan has not significantly changed over the past centuries. Thus, this prior reflects our beliefs that birth year does not heavily impact survival time. For the priors on $\beta_{2j}$, which are coefficients for each of the 4 levels of $type$ (Chinese Emperor, Dalai Lama, Japanese Emperor, and U.S. President), we observed that each leader type is tied to a region. Our research on lifespans across these various regions [[2]][Bibliography] indicates that they are roughly comprable, so we chose a $N(0,1)$ prior for each $\beta_{2j}$. Similarly, we chose a $N(0,1)$ prior for the interaction coefficients $\beta_{3j}$. Here, we did not have a strong enough belief that the coefficients were non-zero and thus chose the same priors as before. 

Our prior on $r$ was informed by the context of the study. Values of $r$ between 0 and 1 indicate a decreasing rate of failure over time, while those greater than 1 represent an increasing rate of failure. Because it can be assumed that the rate of death for world leaders increases with age, we want an $r$ value greater than 1. This belief warranted an informative $\Gamma(10, 0.5)$ prior with mean 5 and variance 2.5, reflecting a relative degree of certainty that $r$ takes on value greater than 1. We perform sensitivity analysis, along with other model diagnostics, in Section 4.2.2. 

# Results 

## Summary of Parameters

```{r}
# The following code generates a table output for the summary of the parameter and coefficient values
post_summary_df <- Bayesian_Leaders_output$BUGSoutput$summary %>% 
  as.data.frame() 
rows <- c("r", "beta_0", "beta_1", 
          "beta_type[1]", "beta_type[2]", "beta_type[3]", "beta_type[4]", "beta_type[5]",
          "beta_int[1]", "beta_int[2]", "beta_int[3]", "beta_int[4]", "beta_int[5]"
)
post_summary_df <- post_summary_df[rows,]
post_summary_df <- post_summary_df %>% 
  tibble::rownames_to_column("Parameter") %>% 
  mutate_if(is.numeric, round, digits = 3) %>% 
  mutate("95% CI" = paste0("(", `2.5%`, ", ", `97.5%`,  ")")) %>% 
  select(-(`2.5%`:`97.5%`)) %>% 
  select(Parameter, mean, sd, `95% CI`) 
post_summary_df$Parameter <- c("r", "(Intercept)", "Time", "Pope", "ChinaEmp", "DalaiLama", "JapanEmp", "UsPres",
                               "Pope*Time", "ChinaEmp*Time", "DalaiLama*Time", "JapanEmp*Time", "UsPres*Time")

post_summary_df %>% 
  mutate(mean = ifelse((mean == 0 & sd == 0), "(Reference)", mean),
         sd = ifelse((mean == "(Reference)" & sd == 0),"", sd),
         `95% CI` = ifelse((mean == "(Reference)" & sd == ""), "", `95% CI`)#,
         #Rhat = ifelse((mean == "(Reference)" & sd == ""), "", Rhat)
  ) %>% 
  kableExtra::kable(booktabs = T, 
                    caption = "Parameters obtained with Model 1, with 95\\% Credible Intervals (CI)",
                    col.names = c("Parameter", "Mean", "Std. Dev.", "95% CI"#, "Rhat"
                    )) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  kableExtra::pack_rows("Type", 4,8) %>% 
  kableExtra::pack_rows("Type*Time", 9,13) %>% 
  row_spec(0, bold = TRUE)
```

Using JAGS, we ran Model 1 with 500,000 simulations to determine the posterior density \newline
$\pi(r,\beta_0,\beta_1, \beta_{2j},\beta_{3j}\, | \,data)$, with results summarized in Table 1. The 95% credible interval is created using the 2.5% and 97.5% quantiles from the posterior distribution. 

## Model Diagnostics 

### Convergence & Autocorrelation Checks 

The traceplots, lag-1 autocorrelation plots, as well as the $\hat{R}$ and $n_{eff}$ values, can be found in the Appendix.

To determine if our model has converged sufficiently, we first examine traceplots for each of the parameters. Traceplots with random scatter around a mean value indicates that the model has converged. $\hat{R}$ values provide an estimate of convergence based on the variance of an estimated parameter between chains, and the variance within a chain. The closer the $\hat{R}$ value to 1, the better the model has converged. Traceplots show the chains have mixed well and the $\hat{R}$ values are all below 1.05, indicating that our model has sufficiently converged. 

To determine if our model has reached a stationary distribution, we first looked at autocorrelation plots. The lag-1 autocorrelation plots peak at the beginning and then bounce around 0. In addition, $n_{eff}$ gives a measure of effective sample size. Ideally, we want this number to equal the number of posterior draws, or at least 200. The autocorrelation plots and large $n_{eff}$ values indicate that our parameters have sufficiently reached stationarity.

### Sensitivity Analysis 

In Figures 7-22 in the Appendix, we explore the posterior distributions of our $r$ parameter, coefficients, and predictions resulting from various priors, where the red lines indicate our posterior mean estimate from our final models. The posterior distributions are roughly the same for the four different $r$ priors, each paired with two different priors on the $\beta$ coefficients. The final plot also compares the posterior distribution of our $r$ and the *time* coefficient. 

### Standardized Residual Plots 

```{r }
# The following code calculates the standardized residuals from our final model
modelOutput <- Bayesian_Leaders_output$BUGSoutput$summary %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column()
y_pred_cols <- modelOutput  %>% 
  pull(rowname)
y_pred_cols <- y_pred_cols[str_detect(y_pred_cols, "y_pred_censored")]
y_post_pred_cols <- c("survival_Francis", "survival_Obama", "survival_DL", "survival_Naruhito")

survivalPredicted <- modelOutput %>% 
  filter(rowname %in% y_pred_cols | rowname %in% y_post_pred_cols) %>% 
  select(rowname, "mean", "sd", "2.5%", "50%", "97.5%")
#creating dataframe with useful information (including prediction residuals and standardized residuals)
leader_order <- c("DalaiLama", "Pope", "JapanEmp", "UsPres")

alive_leaders <- leaders %>% 
  filter(Name %in% alive) %>% 
  slice(match(leader_order, Type))

leader <- c(alive_leaders %>% pull(Name), filtered_leaders$Name)
censored_value <- fct_c(alive_leaders %>% pull(Censored) %>% as.factor(),
                        as.factor(filtered_leaders$Censored))
actual_survival <- c(alive_leaders %>% pull(Age.Event), filtered_leaders$Age.Event)
time <- c(alive_leaders %>% pull(Birth.Date), filtered_leaders$Birth.Date)
survivalPredicted <- survivalPredicted %>%
  mutate(actual = actual_survival) %>%
  mutate(name = leader) %>%
  mutate(Censored = censored_value) %>%
  mutate(Standardized_Residual = (actual-mean)/sd) %>%
  mutate(Residual = actual-mean) %>%
  mutate(time_of_birth = time)

```

```{r fig.height=3.75, fig.align='center', eval=FALSE}
# The following code plots the standardized residuals
ggplot(data = survivalPredicted %>% mutate(Censored = case_when(
  Censored == 1 ~ "Censored",
  TRUE ~ "Non-Censored"
)), aes(x = time_of_birth, y = Standardized_Residual, color = Censored))  + 
  geom_jitter(alpha = 0.6, height = 0.5, width = 0.75) + 
  labs(title = "Standardized Residuals vs Time",
       x = "Time", y = "Standardized Residuals", caption = "Figure 4",
       color = "") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 12)) + 
  scale_shape_manual(values = c(2,4)) +
  scale_colour_manual(values = c("red", "black")) + 
  scale_fill_manual(name = "", labels = c("Non-Censored", "Censored")) 
```

We also examined the standardized residuals, defined as $std\_r_i = (T_i - E[T_i])/sd[T_i]$, and plot $std\_r_i$ against $time_i$ in Figure 4. The plot shows that the standardized residuals are generally randomly dispersed about 0, where the 11 right-censored observations are marked by a red dot. As expected, we have large negative standardized residuals for the censored observations because their current ages are lower than their mean predicted ages. 

### External Validation 

```{r fig.align='center', fig.width=12}
# The following code plots the results of the expected lifepsans for a British Prime Minister & Chinese government leader
birthdate_sun <- "1866-11-12"
death_sun <- "1925-03-12"
birthdate_WC <- "1874-11-30" 
death_WC <- "1965-01-24"
age_sun <- decimal_date(as.Date(death_sun)) - decimal_date(as.Date(birthdate_sun))
age_WC <- decimal_date(as.Date(death_WC)) - decimal_date(as.Date(birthdate_WC))
survivalWC <- Bayesian_Leaders_output$BUGSoutput$sims.matrix[, "survival_WC"] 
survival_sun <- Bayesian_Leaders_output$BUGSoutput$sims.matrix[, "survival_sun"] 
q_WC <- Bayesian_Leaders_output$BUGSoutput$sims.matrix[, "survival_WC"] %>% quantile(c(0.025, 0.5, 0.975))
q_sun <- Bayesian_Leaders_output$BUGSoutput$sims.matrix[, "survival_sun"] %>% quantile(c(0.025, 0.5, 0.975))
gridExtra::grid.arrange(
  ggplot() + 
    geom_histogram(aes(x = survival_sun, y = ..density..), bins=20) + 
    geom_vline(xintercept = age_sun, color = "blue") + 
    geom_vline(xintercept = q_sun[2], color = "red") + 
    geom_vline(xintercept = q_sun[-2], color = "red", linetype = 2) + 
    #annotate("text", x = age_sun + 15, y = 0.0255, label = paste0("Age at Death: \n", round(age_sun,0))) + 
    labs(title = "Sun Yat-sen (Former President of Republic of China)",
         subtitle = "Age of Death: 58",
         y = "Predictive Probability Density", 
         x = "Lifespan (years)",
         caption = "Figure 5") + 
    theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 18)),
  ggplot() + 
    geom_histogram(aes(x = survivalWC, y = ..density..), bins=20) + 
    geom_vline(xintercept = age_WC , color = "blue") + 
    geom_vline(xintercept = q_WC[2], color = "red") + 
    geom_vline(xintercept = q_WC[-2], color = "red", linetype = 2) + 
    #annotate("text", x = age_WC + 8.25, y = 0.020, label = paste0("Age at Death: \n", round(age_WC,0))) + 
    labs(title = "Winston Churchill (Former UK Prime Minister)",
         subtitle = "Age of Death: 90",
         x = "Lifespan (years)",
         caption = "Figure 6") + 
    theme(axis.title.y = element_blank(),
          plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 18)),
  ncol=2,top = textGrob("Posterior Prediction of Lifespan",gp=gpar(fontsize=20)))
```

To further validate our model, we display posterior distributions for Sun Yat-sen, a Chinese leader, in Figure 5 and Winston Churchill, a UK Prime Minister, in Figure 6. We assigned the $type$ of Sun Yat-sen to be that of Chinese Emperors (2), and the $type$ of Winston Churchill to be that of US Presidents (5), since it the role of a Prime Minister and a President are similar in nature. 

These out-of-sample predictions provide reasonable estimates and capture the true lifespans. The blue line indicates the leader's age at death, while the solid red line indicates the posterior median lifespan with corresponding 95% credible intervals shown by the dotted red lines. Both of the leader's true age at death is captured between the median and the 97.5% quantile, with the posterior predictions for Sun doing better than that of Churchill's. This may be because Churchill lived longer than a majority of US Presidents. 

# Discussion 

## Interpretation of Parameters 

Our model for survival time can be reparameterized in such a way that eases interpretability. Specifically, the log mean of $T_i$ can be written as a linear function of our covariates of interest: 

$$
\begin{aligned}
log(mean[T_i]) =  \alpha_0 + \alpha_1\ time_i + \sum_{j=2}^{5} \alpha_{2j} & \ I(type_i = j) + \sum_{j=2}^{5}  \alpha_{3j}\  I(type_i = j)\ *time_i
\end{aligned}
$$

where $\alpha_k = - \frac{\beta_k}{r}$ for any $k$. When exponentiating both sides of this expression, we observe that an increase in one unit of $time$ results in a multiplicative change of $exp(\alpha_1)$ in the mean survival time for a Pope. We can further compute the *percentage* increase in the mean survival time of a Pope given a unit increase in $time$ by computing $100 * (exp(\alpha_1) − 1)$. A similar computation can be performed to understand the effect of increasing time on mean survival time across leader types. This provides an interpretable quantification of the interaction term, $type*time$ appearing in the expression for $log(\mu)$.

```{r}
# The following code calculates the multiplicative difference in lifespans across leadership type we expect if a decade passed
model <- Bayesian_Leaders_output$BUGS$summary %>% as.data.frame()

#Pope
pope_alpha <- c(model$`2.5%`[2], model$mean[2], model$`97.5%`[2])
pope_multiplicative_difference <- exp(0.1*pope_alpha)
pope_percent_difference <- 100*(pope_multiplicative_difference - 1)


#Chinese Emperors
ce_alpha <- c(model$`2.5%`[2] + model$`2.5%`[3], model$mean[2] + model$mean[3], model$`97.5%`[2] + model$`97.5%`[3])
ce_multiplicative_difference <- exp(0.1*ce_alpha)
ce_percent_difference <- 100*(ce_multiplicative_difference - 1)

#Dalai Lama
dl_alpha <- c(model$`2.5%`[2] + model$`2.5%`[4], model$mean[2] + model$mean[4], model$`97.5%`[2] + model$`97.5%`[4])
dl_multiplicative_difference <- exp(0.1*dl_alpha)
dl_percent_difference <- 100*(dl_multiplicative_difference - 1)

#Japanese Emperor

je_alpha <- c(model$`2.5%`[2] + model$`2.5%`[5], model$mean[2] + model$mean[5], model$`97.5%`[2] + model$`97.5%`[5])
je_multiplicative_difference <- exp(0.1*je_alpha)
je_percent_difference <- 100*(je_multiplicative_difference - 1)

#US Presidents

us_alpha <- c(model$`2.5%`[2] + model$`2.5%`[6], model$mean[2] + model$mean[6], model$`97.5%`[2] + model$`97.5%`[6])
us_multiplicative_difference <- exp(0.1*us_alpha)
us_percent_difference <- 100*(us_multiplicative_difference - 1)


Type <- c("Pope", "Chinese Emperor", "Dalai Lama", "Japanese Emperor", "U.S. President")
Multiplicative_Scale_Factor <- data.frame(Type, t(data.frame(pope_multiplicative_difference, ce_multiplicative_difference, dl_multiplicative_difference, je_multiplicative_difference, us_multiplicative_difference))) 
Percent_Difference <- data.frame(Type, t(data.frame(pope_percent_difference, ce_percent_difference, dl_percent_difference, je_percent_difference, us_percent_difference)))

Percent_Difference %>% 
  tibble::rownames_to_column() %>% 
  mutate_if(is.numeric, round, digits=3) %>% 
  mutate(Mean = X2, 
         `95% CI` = paste0("(", X1, ", ", X3, ")")) %>% 
  select(-(X1:X3), -rowname) %>% 
  kableExtra::kable(booktabs = TRUE,
                    caption = "Multiplicative and Percent Change in Mean Survival Time, from decade increase in Birth Year", 
                    col.names = c("Leadership Type", "Mean", "95% CI")) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  kableExtra::pack_rows("Percent Difference", 1,5) %>% 
  row_spec(0, bold = TRUE) %>% 
  column_spec(column = c(2:3), latex_column_spec = "c") %>%   column_spec(column = c(1), width = "20em")
```

In Table 2, we summarize the percent changes in the mean survival time resulting from a decade increase in birth year for each leader type. Recall that $time$ is a scaled and shifted version of birth year that is measured in centuries. We thus consider the changes resulting from a 0.1 unit increase in $time$ in order to make interprations on a decade scale. As can be seen, the 95% credible intervals for Chinese Emperors, Japanese Emperors, and US Presidents contain the value 0. Thus, for these types of leaders, we see that a decade increase in birth year does not yield a significant change in expected survival time. In contrast, the intervals for Popes and Dalai Lamas do not contain 0. Specifically, for a given decade increase in birth year, Dalai Lamas exhibit a decrease in expected survial time while Popes exhibit an increase. Though the point and interval estimates for these values point toward significance, they are still quite close to 0, signaling that a relationship between birth year and survival time is largely independent of leadership type. 

Interpreting the individual effect of $type$ on survival time is not as straightforward as understanding its interaction with $time$. However, if we choose certain values of $time$ to condition on, we can extract meaningful insights. Once conditioned on $time$, we can quantify the percent difference in mean survival time of world leaders in relation to the baseline level of Popes.

```{r}
# The following code walks through the way we derived predictions to answer the posed questions in the introduction
#range of birth years 
#POPES: 1324-1936
#CHINESE EMPEROR: 1328-1906
#DL: 1391-1935
#JAPANESE EMPEROR: 1328-1933
#US PRESIDENT: 1732-1961

#dates: 1800, 1850, 1900
model <- Bayesian_Leaders_output$BUGS$summary %>% as.data.frame()

dates_of_interest <- c(1800, 1850, 1900)
centered_dates_of_interest <- (dates_of_interest - c(1300, 1300, 1300))/100
centered_dates_of_interest <- centered_dates_of_interest - mean(filtered_leaders$time)
Type <- c("Chinese Emperor", "Dalai Lama", "Japanese Emperor", "U.S. President")

#1800:
time <- centered_dates_of_interest[1]

#Chinese Emperors
ce_ft <- c(model$`2.5%`[7] + time*(model$`2.5%`[3]), model$mean[7] + time*(model$mean[3]), model$`97.5%`[7] + time*(model$`97.5%`[3]))
ce_multiplicative_difference_ft <- exp(ce_ft)
ce_percent_difference_ft <- 100*(ce_multiplicative_difference_ft - 1)

#Dalai Lama
dl_ft <- c(model$`2.5%`[8] + time*(model$`2.5%`[4]), model$mean[8] + time*(model$mean[4]), model$`97.5%`[8] + time*(model$`97.5%`[4]))
dl_multiplicative_difference_ft <- exp(dl_ft)
dl_percent_difference_ft <- 100*(dl_multiplicative_difference_ft - 1)

#Japanese Emperor

je_ft <- c(model$`2.5%`[9] + time*(model$`2.5%`[5]), model$mean[9] + time*(model$mean[5]), model$`97.5%`[9] + time*(model$`97.5%`[5]))
je_multiplicative_difference_ft <- exp(je_ft)
je_percent_difference_ft <- 100*(je_multiplicative_difference_ft - 1)

#US Presidents

us_ft <- c(model$`2.5%`[10] + time*(model$`2.5%`[6]), model$mean[10] + time*(model$mean[6]), model$`97.5%`[10] + time*(model$`97.5%`[6]))
us_multiplicative_difference_ft <- exp(us_ft)
us_percent_difference_ft <- 100*(us_multiplicative_difference_ft - 1)


Percent_Difference_1800 <- data.frame(Type, t(data.frame(ce_percent_difference_ft, dl_percent_difference_ft, je_percent_difference_ft, us_percent_difference_ft))) %>%
  mutate_if(is.numeric, round, digits=3) %>% 
  mutate(Mean = X2) %>%
  mutate(`95% CI` = paste0("(", X1, ", ", X3, ")")) %>%
  select(Type, Mean, `95% CI`)

#1850:
time <- centered_dates_of_interest[2]

#Chinese Emperors
ce_ft <- c(model$`2.5%`[7] + time*(model$`2.5%`[3]), model$mean[7] + time*(model$mean[3]), model$`97.5%`[7] + time*(model$`97.5%`[3]))
ce_multiplicative_difference_ft <- exp(ce_ft)
ce_percent_difference_ft <- 100*(ce_multiplicative_difference_ft - 1)

#Dalai Lama
dl_ft <- c(model$`2.5%`[8] + time*(model$`2.5%`[4]), model$mean[8] + time*(model$mean[4]), model$`97.5%`[8] + time*(model$`97.5%`[4]))
dl_multiplicative_difference_ft <- exp(dl_ft)
dl_percent_difference_ft <- 100*(dl_multiplicative_difference_ft - 1)

#Japanese Emperor

je_ft <- c(model$`2.5%`[9] + time*(model$`2.5%`[5]), model$mean[9] + time*(model$mean[5]), model$`97.5%`[9] + time*(model$`97.5%`[5]))
je_multiplicative_difference_ft <- exp(je_ft)
je_percent_difference_ft <- 100*(je_multiplicative_difference_ft - 1)

#US Presidents

us_ft <- c(model$`2.5%`[10] + time*(model$`2.5%`[6]), model$mean[10] + time*(model$mean[6]), model$`97.5%`[10] + time*(model$`97.5%`[6]))
us_multiplicative_difference_ft <- exp(us_ft)
us_percent_difference_ft <- 100*(us_multiplicative_difference_ft - 1)


Percent_Difference_1850 <- data.frame(Type, t(data.frame(ce_percent_difference_ft, dl_percent_difference_ft, je_percent_difference_ft, us_percent_difference_ft))) %>%
  mutate_if(is.numeric, round, digits=3) %>% 
  mutate(Mean = X2) %>%
  mutate(`95% CI` = paste0("(", X1, ", ", X3, ")")) %>%
  select(Type, Mean, `95% CI`)


#1900:
time <- centered_dates_of_interest[3]

#Chinese Emperors
ce_ft <- c(model$`2.5%`[7] + time*(model$`2.5%`[3]), model$mean[7] + time*(model$mean[3]), model$`97.5%`[7] + time*(model$`97.5%`[3]))
ce_multiplicative_difference_ft <- exp(ce_ft)
ce_percent_difference_ft <- 100*(ce_multiplicative_difference_ft - 1)

#Dalai Lama
dl_ft <- c(model$`2.5%`[8] + time*(model$`2.5%`[4]), model$mean[8] + time*(model$mean[4]), model$`97.5%`[8] + time*(model$`97.5%`[4]))
dl_multiplicative_difference_ft <- exp(dl_ft)
dl_percent_difference_ft <- 100*(dl_multiplicative_difference_ft - 1)

#Japanese Emperor

je_ft <- c(model$`2.5%`[9] + time*(model$`2.5%`[5]), model$mean[9] + time*(model$mean[5]), model$`97.5%`[9] + time*(model$`97.5%`[5]))
je_multiplicative_difference_ft <- exp(je_ft)
je_percent_difference_ft <- 100*(je_multiplicative_difference_ft - 1)

#US Presidents

us_ft <- c(model$`2.5%`[10] + time*(model$`2.5%`[6]), model$mean[10] + time*(model$mean[6]), model$`97.5%`[10] + time*(model$`97.5%`[6]))
us_multiplicative_difference_ft <- exp(us_ft)
us_percent_difference_ft <- 100*(us_multiplicative_difference_ft - 1)


Percent_Difference_1900 <- data.frame(Type, t(data.frame(ce_percent_difference_ft, dl_percent_difference_ft, je_percent_difference_ft, us_percent_difference_ft))) %>%
  mutate_if(is.numeric, round, digits=3) %>% 
  mutate(Mean = X2) %>%
  mutate(`95% CI` = paste0("(", X1, ", ", X3, ")")) %>%
  select(Type, Mean, `95% CI`)


bind_rows(Percent_Difference_1800,
          #Percent_Difference_1850, 
          Percent_Difference_1900) %>% 
  kableExtra::kable(booktabs = TRUE,
                    caption = "Percent Difference in Mean Survival Time of World Leaders Compared to Popes by Birth Year", 
                    col.names = c("Leadership Type", "Mean", "95% CI")) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  kableExtra::pack_rows("1800", 1,4) %>% 
  kableExtra::pack_rows("1900", 5,8) #%>% 
#kableExtra::pack_rows("1900", 9,12) 
```

Table 3 shows, at three different birth years, the expected percent difference in the mean survival time of world leaders as compared to the baseline level of Popes. Across the three birth years, 1800, 1850, and 1900, the expected survial times for Chinese Emperors and Dalai Lamas is significantly lower than that of Popes. For example, a Chinese Emperor born in 1900 is expected to have a survival time that is `r -1 * round(ce_percent_difference_ft[2], 1)`% less than a Pope born in the same year. Similarly, a Dalai Lama born in 1900 is expected to have survival time that is `r -1 *round(dl_percent_difference_ft[2], 1)`% less than a Pope born in the same year. 

Althought the average percent difference in expected survival time for Japanese Emperor and U.S. Presidents as compared to Popes is negative across the three birth years, the corresponding credible intervals contain zero, suggesting that these values are not significant.

## Posterior Predictive Distribution 

```{r}
# The following code displays the plots of our posterior predictive distributions
alive_leaders = leaders %>% 
  filter(Name %in% alive) %>% 
  bind_rows(filtered_leaders %>% 
              mutate(rownum = row_number()) %>% 
              filter(Censored == 1))

al_index = alive_leaders %>% 
  filter(!is.na(rownum)) %>% 
  pull(rownum) %>% 
  paste0("y_pred_censored[", ., "]") %>% 
  as.vector()

al_names = alive_leaders %>% pull(Name) %>% str_remove_all("\\s*\\([^\\)]+\\)")
al_cur_age = alive_leaders %>% pull(Age.Event)
al_type = alive_leaders %>% pull(Type)
alive_leaders_post = Bayesian_Leaders_output$BUGSoutput$summary[c("survival_Francis",
                                                                  "survival_Obama",
                                                                  "survival_DL",
                                                                  "survival_Naruhito",
                                                                  al_index), ] %>%
  as.data.frame() %>% 
  tibble::rownames_to_column() %>% 
  mutate(rowname = al_names,
         cur_age = al_cur_age,
         type = al_type
  ) %>% 
  mutate_if(is.numeric, round, digits=3) 
alive_leaders_order <- alive_leaders_post %>% pull(rowname)
alive_leaders_order <- c(alive_leaders_order[-10],  alive_leaders_order[10])
alive_leaders_post %>% 
  slice(match(alive_leaders_order,rowname)) %>% 
  mutate("95% CI" = paste0("(", `2.5%`,", ", `97.5%`, ")")) %>% 
  select(rowname, type,  cur_age, mean, `50%`, sd, "95% CI") %>% 
  kableExtra::kbl(booktabs = TRUE, caption = "Posterior Predictions for Lifespans of Living Leaders, with 95\\% Credible Intervals (CI)", 
                  col.names = c("Leader", "Type", "Current Age", "Mean", "Median", "Std. Dev.", "95% CI")) %>% 
  kable_styling(position = "center",latex_options = c("hold_position")) %>% 
  row_spec(0, bold = TRUE) %>% 
  column_spec(column = c(3:7), latex_column_spec = "c") %>% 
  #kable_styling(position = "center", latex_options = c("hold_position")) %>% 
  add_header_above(c(" " = 3, "Posterior Values" = 4))  
```

Now we will address the research objective of providing posterior predictions of the lifespans of living leaders. Table 4 displays these point and interval predictions. Since we set the upper bound of the truncated Weibull distribution to 110 years, we expect the upper tails of the 95% credible intervals to be near 110 years. This is the case for all the leaders except the 14th Dalai Lama, whose 97.5% quantile is `r alive_leaders_post %>% filter(rowname == "14th Dalai Lama") %>% pull("97.5%")` years, and Emperor Jiawen, whose 97.5% quantile is `r alive_leaders_post %>% filter(rowname == "Jianwen Emperor") %>% pull("97.5%")` years. As previously discussed, there was a string of early deaths among Dalai Lamas in the 19th century, which may have impacted the posterior predictions for the 14th Dalai Lama. Furthermore, Dalai Lamas represent the smallest group of leadership type ($n = 14$), which limits these predictions. Additionally, Chinese Emperors generally have a lower life expectancy than other world leaders. Chinese Emperor Jiawen also disappeared when he was 24, which accounts for a lower 97.5% quantile compared to the other leaders.

## Applications for Select Leaders

```{r fig.align='center', fig.width=12}
# The following code calculates the answers to the third question posed in the intro by plotting the sampled lifespan values from the posterior
#probability that Obama will live longer thatn Naruhito
survival_DL <- Bayesian_Leaders_output$BUGSoutput$sims.matrix[,"survival_DL"]
survival_Francis <-Bayesian_Leaders_output$BUGSoutput$sims.matrix[,"survival_Francis"]

DL_vs_francis <- survival_DL - survival_Francis
DL_vs_francis_data <- as.data.frame(DL_vs_francis)
survival_Obama <- Bayesian_Leaders_output$BUGSoutput$sims.matrix[,"survival_Obama"]
survival_Naruhito <- Bayesian_Leaders_output$BUGSoutput$sims.matrix[,"survival_Naruhito"]

Obama_vs_naruhito <- survival_Obama - survival_Naruhito
Obama_vs_naruhito_data <- as.data.frame(Obama_vs_naruhito)





gridExtra::grid.arrange(ggplot(DL_vs_francis_data, aes(x = DL_vs_francis, y = ..density..)) + 
                          geom_histogram(binwidth = 4) + 
                          labs(x = "Difference in Life Expectancy (Dalai Lama - Francis)",
                               y = "Predictive Probability Density", 
                               
                               subtitle = "Dalai Lama and Pope Francis",
                               caption = "Figure 7") + 
                          
                          geom_vline(xintercept=mean(DL_vs_francis), color = "blue") + 
                          theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 18)),
                        ggplot(Obama_vs_naruhito_data, aes(x = Obama_vs_naruhito, y = ..density..)) + 
                          geom_histogram(binwidth = 4) + 
                          labs(
                            subtitle = "President Obama and Emperor Naruhito",
                            x = "Difference in Life Expectancy (Obama - Naruhito)", 
                            y = "Predictive Probability Density",
                            caption = "Figure 8")  +  
                          
                          geom_vline(xintercept=mean(Obama_vs_naruhito), color = "blue") + 
                          theme(axis.title.y = element_blank(),
                                plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 18)),
                        ncol = 2,top = textGrob("Distribution of Difference in Predicted Life Expectancy",gp=gpar(fontsize=20)))


```

Figures 7 and 8 show the distribution of difference in predicted life expectancy of select leaders. 

From the posterior predictive samples for the survival times of the 14th Dalai Lama and Pope Francis, we note that only in  `r round(100*mean(survival_DL>survival_Francis),1)`% of the simulated samples is the Dalai Lama predicted to have a longer lifespan. Thus, there is only a `r round(mean(survival_DL>survival_Francis),3)` probability that the 14th Dalai Lama will have a longer lifespan than Pope Francis. Figure 5 indicates that, on average, the posterior predictions for the survival time of the 14th Dalai Lama are `r -1*round(1*mean(DL_vs_francis), 3)` years (shown by the blue line) below that of Pope Francis. 

From the posterior predictive samples for the survival times of President Obama and Emperor Naruhito, we note that in  `r round(100*mean(survival_Obama > survival_Naruhito),1)`% of the simulated samples, Obama is predicted to have a longer lifespan. Consequently, there is a `r round(mean(survival_Obama > survival_Naruhito), 3)` probability that Obama  will have a longer lifespan than Naruhito. Figure 6 indicates that, on average, the posterior predictions for the survival time of President Obama are only `r round(mean(Obama_vs_naruhito, 3))` years above that of Emperor Naruhito. 

# Conclusion & Limitations 

Revisiting our original question of how year of birth and type of leadership are related to a leader’s survival time, our analysis reveals interesting findings. Additionally, having answered the rest of the questions outlined in the introduction, an intuition for understanding the nuances of our results can be explained by some of the limitations in the study. These include, but are not limited to, small sample size (as a whole and by leader type), lack of health-related covariates, underrepresentation of women, unavailability of studies that inform strong prior distributions, and the absence of informative covariates that represent type of death (i.e., the case of the four short-lived, 19th-century Dalai Lamas and the disappearance of Chinese Emperor Jianwen). These limitations greatly undermine the heterogenous notions of “world leaders”; for example, without including health covariates, this limits our empirical knowledge of world leaders and the availability of healthcare in certain areas during a specific time period. We suggest that future studies should explore the lifespans of “unconventional” leaders, who don’t necessarily hold an official position of office, along with established leaders from Latin America, South America, Africa, and Southeast Asia.


--Could potentiall add more here--
Our analysis supports that there is not a significant association between birth year and survival time among various world leaders. Further, our model has found that a leader's survival time is not dependent on their type of leadership. 
Having answered these questions, we will now state the limitations. These include, but are not limited to, small sample size, lack of health-related covariates, underrepresentation of women, and the unavailability of studies that inform strong prior distributions. We suggest that future studies should explore the lifespans of “unconventional” leaders, who don’t necessarily hold an official position of office, along with established leaders from Latin America, Africa, and Southeast Asia.

# Appendix 

## Bibliography

1. Julian Stander, Luciana Dalla Valle & Mario Cortina-Borja (2018) A Bayesian Survival Analysis of a Historical Dataset: How Long Do Popes Live?, The American Statistician, 72:4, 368-375, DOI: 10.1080/00031305.2017.1328374
2. World Population Prospects - Population Division. (n.d.). Retrieved August 31, 2020, from https://population.un.org/wpp/Download/Standard/Population/
3. Supercentenarian Research Foundation. (n.d.). Retrieved August 31, 2020, from http://supercentenarian-research-foundation.org/TableE.aspx
4. Ruggeri, A. (2018, October 2). Do we really live longer than our ancestors? Retrieved August 31, 2020, from https://www.bbc.com/future/article/20181002-how-long-did-ancient-people-live-life-span-versus-longevity
5. DALAI LAMA- A SHORT LIVED POSITION. (n.d.). Retrieved August 31, 2020, from https://www.tripsatasia.com/newsletter-articles/dalai-lama-a-short-lived-position

```{r}
knitr::opts_chunk$set(eval = FALSE,
                      message = FALSE,
                      warning = FALSE)
```


## Traceplots and Autocorrelation Plots

```{r eval=TRUE, fig.height=4, fig.align='center'}
# The following code chunks plots the convergence of the various parameters & coeffs in this model
tidybayes::gather_draws(Bayesian_Leaders_output %>% coda::as.mcmc(), r,  beta_0, beta_1) %>% 
  ungroup() %>% 
  # tidyr::gather(key, value, "beta_0", "beta_1", "r", -i) %>% 
  ggplot() + 
  geom_line(aes(x = .iteration, y = .value, color = as.factor(.chain)), alpha = 0.7) + 
  facet_grid(.variable~., scales = "free_y") + 
  labs(title = expression(paste("Traceplots of ",beta[0], ", ", beta[1], ", and r")), 
       x = "Iteration", y = "Value",  color = "Chain")
```


```{r eval=TRUE, fig.height=4, fig.align='center'}
tidybayes::gather_draws(Bayesian_Leaders_output %>% coda::as.mcmc(), beta_type[i]) %>%
  ungroup() %>%
  filter(i > 1) %>%
  mutate(.variable = ifelse(is.na(i), .variable, paste0(.variable, "[",i, "]"))) %>%
  # tidyr::gather(key, value, "beta_0", "beta_1", "r", -i) %>%
  ggplot() +
  geom_line(aes(x = .iteration, y = .value, color = as.factor(.chain)), alpha = 0.7) +
  facet_grid(.variable~., scales = "free_y") +
  labs(title = expression(paste("Traceplots of ", beta["2j"], " (coefficients for Type)")),
       x = "Iteration", y = "Value",  color = "Chain")
```


```{r eval=TRUE, fig.height=4, fig.align='center'}
tidybayes::gather_draws(Bayesian_Leaders_output %>% coda::as.mcmc(), beta_int[i]) %>% 
  ungroup() %>% 
  filter(i > 1) %>% 
  mutate(.variable = ifelse(is.na(i), .variable, paste0(.variable, "[",i, "]"))) %>% 
  # tidyr::gather(key, value, "beta_0", "beta_1", "r", -i) %>% 
  ggplot() + 
  geom_line(aes(x = .iteration, y = .value, color = as.factor(.chain)), alpha = 0.7) + 
  facet_grid(.variable~., scales = "free_y") + 
  labs(title = expression(paste("Traceplots of ", beta["3j"], " (coefficients for Type * Time)")), 
       x = "Iteration", y = "Value",  color = "Chain")
```


```{r eval=TRUE, fig.width=3}
# The following function plots the lag 1 autocorrelation for the input paramater
plot_acf <- function(parameter) {
  Bayesian_Leaders_output$BUGSoutput$sims.matrix %>%
    as_tibble() %>%
    pull(parameter) %>%
    acf(main = paste0("Lag-1 Autocorrelation for ", parameter))
}
for (param in c("beta_0", "beta_1")) {
  plot_acf(param)
}
```

## $\hat{R}$ and $n_{eff}$

Note that $n_{eff}$ values for $\beta_{21}$ and $\beta_{31}$ are 1, because forced these parameters to be 0 to reflect the baseline category of $type$. 
```{r eval=TRUE}
# The following displays the effective sample size values for the coeffs and parameters in this model, along with the Rhat values
rhat_neff = Bayesian_Leaders_output$BUGSoutput$summary %>% as.data.frame() %>% select(Rhat, `n.eff`) 
rows <- c("r", "beta_0", "beta_1", 
          "beta_type[1]", "beta_type[2]", "beta_type[3]", "beta_type[4]", "beta_type[5]",
          "beta_int[1]", "beta_int[2]", "beta_int[3]", "beta_int[4]", "beta_int[5]"
)
rhat_neff[rows,] %>% 
  tibble::rownames_to_column() %>%
  kableExtra::kbl(booktabs = TRUE, 
                  col.names = c("Parameter", "Rhat", "N_Eff"))%>% 
  kable_styling(position = "center")
```

## Sensitivity Analysis: Parameters, Coefficients, Predictions Over r and Beta Priors

```{r eval=TRUE}
# The following code generates a table of posterior distributions of model coeffs and parameters across the different priors examined
r_vals <- c("1e-3", "1", "10_5e-1", "25_2e-1")
b_vals <- c("1", "100")
sensitivity_plot = function(param, renamed_param = NA) {
  df = data.frame()
  for (curr_r in r_vals) {
    for (curr_b in b_vals) {
      file_name = paste0("leaders_model_r_",curr_r,"_beta_",curr_b,".Rdata")
      load(file_name)
      
      curr_param = param
      rjags.df <- rjags_model$BUGSoutput$sims.matrix %>% as.data.frame()
      if (!is.na(renamed_param)){
        colnames(rjags.df)[colnames(rjags.df) == param] = renamed_param
        curr_param = renamed_param
      }
      
      param_df = rjags.df %>%
        select(curr_param) %>% 
        mutate(curr_r = case_when(
          curr_r == "1e-3" ~ "0.001",
          curr_r == "10_5e-1" ~ "10_0.5",
          curr_r == "25_2e-1" ~ "25_0.2",
          TRUE ~ as.character(curr_r)),
          priors = ifelse(str_detect(curr_r, "_"),
                          paste0("r~Ga(", str_split(curr_r, "_")[[1]][1], ",", str_split(curr_r, "_")[[1]][2], "),B~N(0,", curr_b, ")"),
                          
                          paste0("r~Exp(", curr_r, "),B~N(0,", curr_b, ")")
          ))
      df = df %>% bind_rows(param_df)
    }
  }
  return(df) 
}
``` 

```{r eval=TRUE}
# The following code chunks display the sensitivity of the input parameter relative to other priors we examined
mean_vals <- Bayesian_Leaders_output$BUGSoutput$summary %>% as.data.frame() %>% select(mean)

get_mean_val <- function(param, renamed_param=NA){
  if (!is.na(renamed_param)){
    mean_vals.df <- as.data.frame(mean_vals)
    rownames(mean_vals.df)[rownames(mean_vals.df) == param] = renamed_param
    mean_vals <- as.matrix(mean_vals.df)
    param = renamed_param
  }
  return(c(mean_vals[param,]))
}

predicted_r <- get_mean_val("r")
predicted_beta_1 <- get_mean_val("beta_1")
predicted_dl <- get_mean_val("survival_DL")
predicted_francis <- get_mean_val("survival_Francis")
predicted_naruhito <- get_mean_val("survival_Naruhito")
predicted_obama <- get_mean_val("survival_Obama")

predicted_ChinaEmp <- get_mean_val("beta_int[2]","ChinaEmp")
predicted_DalaiLama <- get_mean_val("beta_int[3]","DalaiLama")
predicted_JapanEmp <- get_mean_val("beta_int[4]","JapanEmp")
predicted_USPres <- get_mean_val("beta_int[5]","USPres")

predicted_ChinaEmpTime <- get_mean_val("beta_type[2]","ChinaEmpTime")
predicted_DalaiLamaTime <- get_mean_val("beta_type[3]","DalaiLamaTime")
predicted_JapanEmpTime <- get_mean_val("beta_type[4]","JapanEmpTime")
predicted_USPresTime <- get_mean_val("beta_type[5]","USPresTime")
```


```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("r")  %>% 
  ggplot() + 
  geom_density(aes(x = r)) + geom_vline(xintercept = predicted_r, color = "red") + 
  labs(title = expression(paste("r : Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 7") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2) 
```

While the $r$ parameter appears to be slightly more sensitive to priors, the models that it appears to differ for are used highly informative priors on r. This may have constrained our sampler from properly exploring the space. We can see evidence for this when we examine the two plots with the highly non-informative prior $r \sim exp(0.001)$, which align most strongly with our chosen prior  $r \sim \Gamma(10,0.5)$. This is true also for the Dalai Lama coefficient and Dalai Lama * Time coefficient.

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("beta_1")  %>%
  ggplot() +
  geom_density(aes(x = beta_1)) + geom_vline(xintercept = predicted_beta_1, color = "red") +
  labs(title = expression(paste("Time Coefficient: Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 9") +
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) +
  facet_wrap(priors~.,  scales = "free_y", nrow = 2)
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("survival_DL")  %>%
  ggplot() +
  geom_density(aes(x = survival_DL)) + geom_vline(xintercept = predicted_dl, color = "red") +
  labs(title = expression(paste("14th Dalai Lama: Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 10") + xlim(80,90) + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) +
  facet_wrap(priors~.,  scales = "free_y", nrow = 2) 
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("survival_Francis")  %>%
  ggplot() +
  geom_density(aes(x = survival_Francis)) + geom_vline(xintercept = predicted_francis, color = "red") +
  labs(title = expression(paste("Pope Francis: Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 11") + xlim(88,100) + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) +
  facet_wrap(priors~.,  scales = "free_y", nrow = 2)
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("survival_Naruhito")  %>%
  ggplot() +
  geom_density(aes(x = survival_Naruhito)) + geom_vline(xintercept = predicted_naruhito, color = "red") +
  labs(title = expression(paste("Emperor Naruhito: Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 12") + xlim(70,80) + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) +
  facet_wrap(priors~.,  scales = "free_y", nrow = 2)
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("survival_Obama")  %>%
  ggplot() +
  geom_density(aes(x = survival_Obama)) + geom_vline(xintercept = predicted_obama, color = "red") +
  labs(title = expression(paste("President Obama: Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 13") + xlim(75,85) + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) +
  facet_wrap(priors~.,  scales = "free_y", nrow = 2)
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("beta_int[2]","ChinaEmp")  %>% 
  ggplot() + 
  geom_density(aes(x = ChinaEmp)) + geom_vline(xintercept = predicted_ChinaEmp, color = "red") + 
  labs(title = expression(paste("Chinese Emperor Coefficient : Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 14") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2)
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("beta_int[3]","DalaiLama")  %>% 
  ggplot() + 
  geom_density(aes(x = DalaiLama)) + geom_vline(xintercept = predicted_DalaiLama, color = "red") + 
  labs(title = expression(paste("Dalai Lama Coefficient : Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 15") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2)
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("beta_int[4]","JapanEmp")  %>% 
  ggplot() + 
  geom_density(aes(x = JapanEmp)) + geom_vline(xintercept = predicted_JapanEmp, color = "red") + 
  labs(title = expression(paste("Japanese Emperor Coefficient : Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 16") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2) 
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("beta_int[5]","USPres")  %>% 
  ggplot() + 
  geom_density(aes(x = USPres)) + geom_vline(xintercept = predicted_USPres, color = "red") + 
  labs(title = expression(paste("US President * Time Coefficient : Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 17") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2) 
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("beta_type[2]","ChinaEmpTime")  %>% 
  ggplot() + 
  geom_density(aes(x = ChinaEmpTime)) + geom_vline(xintercept = predicted_ChinaEmpTime, color = "red") + 
  labs(title = expression(paste("Chinese Emperor * Time Coefficient : Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 18") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2)
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("beta_type[3]","DalaiLamaTime")  %>% 
  ggplot() + 
  geom_density(aes(x = DalaiLamaTime)) + geom_vline(xintercept = predicted_DalaiLamaTime, color = "red") + 
  labs(title = expression(paste("Dalai Lama * Time Coefficient: Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 19") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2)
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("beta_type[4]","JapanEmpTime")  %>% 
  ggplot() + 
  geom_density(aes(x = JapanEmpTime)) + geom_vline(xintercept = predicted_JapanEmpTime, color = "red") + 
  labs(title = expression(paste("Japanese Emperor * Time Coefficient : Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 20") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2) 
```

```{r eval=TRUE, fig.width=12, fig.align='center'}
sensitivity_plot("beta_type[5]","USPresTime")  %>% 
  ggplot() + 
  geom_density(aes(x = USPresTime)) + geom_vline(xintercept = predicted_USPresTime, color = "red") + 
  labs(title = expression(paste("US President * Time Coefficient : Sensitivity Analysis Over r and ", beta)),
       x = expression(beta[1]),caption="Figure 21") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2) 
```


## Sensitivity Analysis: r, Time Coefficient Over Different Upper Bounds

```{r eval=TRUE}
upper_vals <- c("100","105","")

sensitivity_plot_trunc = function(param, renamed_param = NA) {
  df = data.frame()
  for (curr_upper in upper_vals) {
    upperText = paste0("_upper_",curr_upper)
    if (curr_upper == ""){
      upperText = ""
    }
    file_name = paste0("leaders_model_r_10_5e-1_beta_1",upperText,".Rdata")
    load(file_name)
    
    curr_param = param
    rjags.df <- rjags_model$BUGSoutput$sims.matrix %>% as.data.frame()
    if (!is.na(renamed_param)){
      colnames(rjags.df)[colnames(rjags.df) == param] = renamed_param
      curr_param = renamed_param
    }
    
    param_df = rjags.df %>%
      select(curr_param) %>% 
      mutate(curr_upper = case_when(
        curr_upper == "" ~ "110",
        TRUE ~ as.character(curr_upper)),
        priors = paste0("Upper Bound = ",curr_upper," years"))
    df = df %>% bind_rows(param_df)
  }
  return(df) 
}

sensitivity_plot_trunc("r")  %>% 
  ggplot() + 
  geom_density(aes(x = r)) + geom_vline(xintercept = predicted_r, color = "red") + 
  labs(title = expression(paste("r : Sensitivity Analysis Over 100, 105, 110 Years as Upper Bounds")),
       x = expression(beta[1]),caption="Figure 21") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2) 
```

```{r eval=TRUE}
sensitivity_plot_trunc("beta_1")  %>% 
  ggplot() + 
  geom_density(aes(x = beta_1)) + geom_vline(xintercept = predicted_beta_1, color = "red") + 
  labs(title = expression(paste("Time Coefficient : Sensitivity Analysis Over 100, 105, 110 Years as Upper Bounds")),
       x = expression(beta[1]),caption="Figure 22") + 
  theme(plot.caption = element_text(hjust = 0.5, vjust = -0.5, size = 16)) + 
  facet_wrap(priors~.,  scales = "free_y", nrow = 2) 
```

## Calculations from Section 5.1

Below, we obtain the expected multiplicative factor of being a certain leader type of mean birth year compared to being a Pope of mean birth year.

```{r echo=TRUE}
# The following code shows how we obtained the multiplicative factor of being a certain leader type of mean birth year, relative to the Pope
model <- Bayesian_Leaders_output$BUGS$summary %>% as.data.frame()

#Chinese Emperors
ce_int <- exp(model$mean[7])

#Dalai Lamas
dl_int <- exp(model$mean[8])

#Japanese Emperor
je_int <- exp(model$mean[9])

#US Pres
us_int <- exp(model$mean[10])
model <- Bayesian_Leaders_output$BUGS$summary %>% as.data.frame()
#Chinese Emperors
ce_int <- exp(model$mean[7])
#Dalai Lamas
dl_int <- exp(model$mean[8])
#Japanese Emperor
je_int <- exp(model$mean[9])
#US Pres
us_int <- exp(model$mean[10])
```

